---
title: 'Milestone #2'
author: "Virginia Chan, Patrick Traynor, Antoinette Stetzenmeyer"
date: "9/29/2021"
output: pdf_document
---



```{r}
#load libraries
library(tidyverse)
library(lubridate)
```

Description of dataset
What is the data source? (1-2 sentences on where the data is coming from, dates included, etc.)

The data source for the COVID-19 Vaccine Administered by Zip Code dataset is the California Immunization Registry and the American Community Survey's 2015-2019 5-Year data. 

The data source for the CA County Demographic dataset is based on US census data.

How does the dataset relate to the group problem statement and question?

There is growing concern that counties with younger median age are less likely to have higher counts of vaccinated persons. To investigate this, we need to explore the CA census demographics to compare median age values and proportions of vaccinate persons together on a county level. This will help us determine if there is any relationship between those two variables. 

\newpage 

```{r}
#Import statement
library(readr)
cov_vax_admin <- read_csv("cov_vax_admin.csv")
cov_vax_admin

ca_county_demographics <- read_csv("ca_county_demographics.csv")
ca_county_demographics
```

\newpage 

Identify data types for 5+ data elements/columns/variables for both data sets

```{r}
#2 Identify data types for 5+ data elements/columns/variables

#viewing structure of both datasets

str(ca_county_demographics)

str(cov_vax_admin)
```
```

The 5+ data elements/columns/variables we are interested in: 

From COVID vax data set: county (county: chr), persons fully vaxxed (persons_fully_vaccinated: num), age 12+ population (age12_plus_population: num), zip code (zip_code_tabulation_area: num).

County is the only character value of interest while the rest of the variables are numeric. 

From the Census data set: median age (med_age: num), county name (name: chr), and overall population (pop2012: num). 

County name is the only character value of interest while the rest of the variables are numeric. 

All the variables we are interested in are in the desired type/format and we will not need to
convert any columns to numeric or another type. 

\newpage 





#3 Provide a basic description of the 5+ data elements
Numeric: mean, median, range
Character:  unique values/categories
Or any other descriptives that will be useful to the analysis


```{r}
ca_county_demographic_total_mean_age <- ca_county_demographics %>%
  summarize(total_mean_age = mean(med_age, na.rm = TRUE)) 
  
(view(ca_county_demographic_total_mean_age))

# A tibble: 1 x 1
  #total_mean_age
          # <dbl>
#1           38.5

#This is the unweighted mean of the medium ages across each county in 
#California

#Exploring distinct county and zipcodes in California

unique(cov_vax_admin$county)

unique(cov_vax_admin$zip_code_tabulation_area)

#Exploring the total number of persons totally vaccinated per County

total_vaccinated_california <- cov_vax_admin %>%
group_by(county)%>%
summarize(total_vax = sum(persons_fully_vaccinated,na.rm = TRUE))
  
```
example

flights_mean_delay <- flights %>%
  group_by(month, carrier) %>%
  summarize(total_flights = n_distinct(flight), delay_avg = mean(dep_delay, na.rm = TRUE)) 