---
title: 'Milestone #2'
author: "Virginia Chan, Patrick Traynor, Antoinette Stetzenmeyer"
date: "9/29/2021"
output: pdf_document
---



```{r}
#load libraries
library(tidyverse)
library(lubridate)
```

Description of dataset
What is the data source? (1-2 sentences on where the data is coming from, dates included, etc.)

The data source for the COVID-19 Vaccine Administered by Zip Code dataset is the California Immunization Registry and the American Community Survey's 2015-2019 5-Year data. 

The data source for the CA County Demographic dataset is based on US census data.

How does the dataset relate to the group problem statement and question?

There is growing concern that counties with younger median age are less likely to have higher counts of vaccinated persons. To investigate this, we need to explore the CA census demographics to compare median age values and proportions of vaccinate persons together on a county level. This will help us determine if there is any relationship between those two variables. 

\newpage 

```{r}
#Import statement
library(readr)
cov_vax_admin <- read_csv("cov_vax_admin.csv")
cov_vax_admin

ca_county_demographics <- read_csv("ca_county_demographics.csv")
ca_county_demographics
```

\newpage 

Identify data types for 5+ data elements/columns/variables for both data sets

```{r}
#2 identify data types for 5+ data elements/columns/variables

str(ca_county_demographics)
```
```





explore some CA census demographics to compare median age values and proportions of vaccinated persons together on a county level

Columns/rows interested in: name of county, pop12 sqmi, median age, make up of 


From COVID vax data set: count, persons fully vaxxed, persons partially vaxxed, age 12+ population(?)







```{r}

#identify data types for 5+ data elements/columns/variables

str(cov_vax_admin)
```
want to know proportion of individuals vaccinated by age, county, and local health jurisdiciton.
Some columns: as_of_date, zip code tabluation area, county, vaccine equity metric quartile, age 12 plus population, persons fully vaccinated, persons partially vaccinated, vem source (healthy places index score)

Identify the desired type/format for each variable—will you need to convert any columns to numeric or another type? 



\newpage 


#3 Provide a basic description of the 5+ data elements
Numeric: mean, median, range
Character:  unique values/categories
Or any other descriptives that will be useful to the analysis


```{r}
ca_county_demographic_total_median_age <- ca_county_demographics %>%
  summarize(total_median_age = mean(med_age, na.rm = TRUE)) 

  ca_county_demographic_range_age <- ca_county_demographics %>%
  summarize(range = range(med_age, na.rm = TRUE)) 
  
  
```
example

flights_mean_delay <- flights %>%
  group_by(month, carrier) %>%
  summarize(total_flights = n_distinct(flight), delay_avg = mean(dep_delay, na.rm = TRUE)) 